Untitled

### 2021-03-14 analysis/chapters/paper_ana_20181122.Rmd
Init paper | Dominio

---
title: "FCST"
output: 
  bookdown::html_document2: default
  bookdown::pdf_document2: default
bibliography: mesoda_ref.bib
---

@@

### FSS

Acumulados a 6 hs

```{r}
files <- Sys.glob("/home/paola.corrales/mesoda/analysis/data/derived_data/fss_6h_fcst_ens_*_E[4-7]_ens.csv")

fss <- purrr::map(files, function(f) {

  meta <- unglue(basename(f), "fss_6h_fcst_ens_{ini_date}_{exp}_ens.csv")

  fread(f) %>% 
    .[, date := as_datetime(date)] %>% 
    .[, ini_date := ymd_h(meta[[1]][["ini_date"]])]
}) %>% 
  rbindlist()

fss %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp, linetype = factor(ini_date))) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date(20181122000000, 20181123000000, "12 hours") +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("1" = "1 km", "5" = "50 km", 
                                            "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, linetype = NULL) +
  theme_minimal()

### Soundings

```{r message=FALSE, warning=FALSE}
# files <- Sys.glob("/home/paola.corrales/datosmunin/EXP/E*/FCST/*/sondeos/*")
# files <- files[str_detect(files, "_01_")]
# 
# sondeos <- purrr::map(files, function(f){
#   
#   fread(f) %>% 
#     .[, value := fifelse(variable %in% c("t", "td"), value + 273.15, value)] %>% 
#     .[, launch_time := as_datetime(launch_time)] %>% 
#     .[, fcst_exp := basename(dirname(dirname(f)))]
#   
# }) %>% 
#   rbindlist()

miembros <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_agregados.rds") %>% 
  .[, mem := formatC(mem, width = 3, flag = "0")] %>% 
  fisica[., on = "mem"] %>% 
  separate(fisica, into = c("cumulus", "pbl"))

media <- read_rds("/home/paola.corrales/datosmunin/EXP/derived_data/sondeos_media.rds") %>% 
  .[, .(RMSE = mean(sqrt((value - fcst_value)^2), na.rm = TRUE), 
        BIAS = mean(value - fcst_value, na.rm = TRUE)), 
    by = .(iop, exp, variable, lev, fcst_exp)] %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico", value.name = "value.est")

IOP <- tribble(
  ~iop, ~ini, ~end,
  "IOP07", ymd_hms("20181121150000"), ymd_hms("20181121210000"),
  "IOP08", ymd_hms("20181122140000"), ymd_hms("20181122200000")
) %>% setDT()

```{r}

media %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-ensamble}
extremos <- miembros %>% 
  .[, .(max_rmse = max(RMSE),
        min_rmse = min(RMSE)), by = .(exp, fcst_exp, lev, variable)]

media %>% 
  .[estadistico == "RMSE"] %>% 
  extremos[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_ribbon(aes(ymin = min_rmse, 
                  ymax = max_rmse, fill = exp), alpha = 0.2) +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E7 = "RAD")) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

media %>% 
  .[estadistico == "RMSE"] %>% 
  miembros[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(alpha = 0.1, 
            aes(y = RMSE, color = exp, group = interaction(exp, mem))) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()
```

```{r rmse-sondeos-param}
pbl <- miembros[, .(RMSE = mean(RMSE),
                    tipo = "pbl"), 
                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
  setnames("pbl", "param")

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           RMSE = value.est, 
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(RMSE = mean(RMSE),
                    tipo = "cumulus"), 
                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
  setnames("cumulus", "param")

media %>% 
  .[estadistico == "RMSE"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "RMSE" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           RMSE = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, RMSE)) +
  # geom_hline(yintercept = 0, color = "grey30") +
  geom_line(aes(y = RMSE, color = exp, linetype = param, group = interaction(exp, param))) +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

rbind(pbl, cum) %>% 
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp)] %>% 
  ggplot(aes(RMSE_mean)) +
  geom_histogram(aes(fill = exp), 
                 position = "identity", 
                 alpha = 0.5) +
   scale_fill_manual(values = colores_exp, 
                     labels = c(E4 = "CONV", E5 = "AUT",
                                E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x")

rbind(pbl, cum) %>%
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(RMSE_mean, exp)) +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x") +
  labs(y = NULL, x = "mean RMSE", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))

```{r bias-sondeos-ensamble}
extremos <- miembros %>% 
  .[, .(max_bias = max(BIAS),
        min_bias = min(BIAS)), by = .(exp, fcst_exp, lev, variable)]

media %>% 
  .[estadistico == "BIAS"] %>% 
  extremos[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +
  geom_hline(yintercept = 0, color = "grey30") +
  geom_ribbon(aes(ymin = min_bias, 
                  ymax = max_bias, fill = exp), alpha = 0.2) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E7 = "RAD")) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

media %>% 
  .[estadistico == "BIAS"] %>% 
  miembros[., on = .NATURAL] %>% 
  ggplot(aes(lev/1000, value.est)) +

  geom_line(alpha = 0.1, 
            aes(y = BIAS, color = exp, group = interaction(exp, mem))) +
  geom_line(aes(color = exp, linetype = estadistico)) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp + iop~variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal() 
```

```{r bias-sondeos-param}
pbl <- miembros[, .(BIAS = mean(BIAS),
                    tipo = "pbl"), 
                by = .(exp, fcst_exp, pbl, lev, variable)] %>% 
  setnames("pbl", "param")

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           BIAS = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

cum <- miembros[, .(BIAS = mean(BIAS),
                    tipo = "cumulus"), 
                by = .(exp, fcst_exp, cumulus, lev, variable)] %>% 
  setnames("cumulus", "param")

media %>% 
  .[estadistico == "BIAS"] %>% 
  .[, c("exp", "variable", "lev", "fcst_exp", "BIAS" = "value.est")] %>% 
  .[, ":="(param = "Mean",
           BIAS = value.est,
           tipo = "Mean")] %>% 
  .[, value.est := NULL] %>% 
  rbind(., cum) %>% 
  ggplot(aes(lev/1000, BIAS)) +
  geom_line(aes(y = BIAS, color = exp, linetype = param, group = interaction(exp, param))) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  coord_flip() +
  facet_grid(fcst_exp ~ variable, scales = "free_x", 
             labeller = labeller(variable = c("t" = "Temperature",
                                              "td" = "Dew point\ntemperature",
                                              "u" = "U wind",
                                              "v" = "V wind"))) +
  labs(y = NULL, x = "Height (Km)", color = NULL, linetype = NULL) +
  theme_minimal()

rbind(pbl, cum) %>%
  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
    by = .(variable, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(BIAS_mean, exp)) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ variable, scales = "free_x") +
  labs(y = NULL, x = "mean BIAS", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))

### Observaciones de superficie

```{r}
# files <- Sys.glob(paste0(derived_data, "/interp_obs/2018112200/interp*"))
# 
# obs <- purrr::map(files, function(f) {
#   print(f)
#   read_rds(f) %>%
#     .[, id := 1:.N, by = ens] %>%
#     .[, fcst_exp := "2018112200"] %>%
#     .[, exp := NULL] %>%
#     melt(measure.vars = c("E4", "E5", "E6", "E7"),
#          variable.name = "exp",
#          value.name = "fcst") %>%
#     .[, ":="(obs = if_else(usage.flag == 1, obs, NA_real_),
#              fcst = if_else(usage.flag == 1, fcst, NA_real_))] %>%
#     .[, c(list(obs = mean(obs, na.rm = TRUE),
#                fcst = mean(fcst, na.rm = TRUE)),
#           .SD[ens == 1]), by = .(id, exp), .SDcols = -c("obs", "fcst")] %>%
#     .[usage.flag == 1]
# 
# }) %>%
#   rbindlist()

obs_00 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean.rds"))
obs_06 <- read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean.rds"))

media <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mean2.rds")),
               read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mean2.rds"))) %>% 
  .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
           BIAS = fifelse(var == "q", BIAS*1000, BIAS))]

miembros <- rbind(read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112200_mem.rds")),
                  read_rds(paste0(derived_data, "/interp_obs/interp_obs_2018112206_mem.rds"))) %>% 
    .[, ":="(RMSE = fifelse(var == "q", RMSE*1000, RMSE),
           BIAS = fifelse(var == "q", BIAS*1000, BIAS))]

media %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico") %>% 
  .[estadistico == "RMSE"] %>% 
  ggplot(aes(date, value)) +
  geom_line(aes(color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()

media %>% 
  melt(measure.vars = c("RMSE", "BIAS"), variable.name = "estadistico") %>% 
  .[estadistico == "BIAS"] %>% 
  ggplot(aes(date, value)) +
  geom_hline(yintercept = 0, color = "grey60") +
  geom_line(aes(color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()
```

```{r}

extremos <- miembros %>% 
  .[, .(max_rmse = max(RMSE),
        min_rmse = min(RMSE)), by = .(var, exp, fcst_exp, date)]

media %>% 
  extremos[., on = .NATURAL] %>% 
  .[exp == "E7"] %>% 
  ggplot(aes(date, RMSE)) +
  geom_ribbon(aes(ymin = min_rmse, 
                  ymax = max_rmse, fill = exp, group = interaction(exp, fcst_exp)), alpha = 0.2) +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E7 = "RAD")) +
  geom_line(aes(color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()

media %>% 
  .[, mem := "000"] %>% 
  rbind(., miembros) %>% 
  .[exp == "E7"] %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(data = ~.x[mem != "000"], alpha = 0.1, 
            aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
  geom_line(data = ~.x[mem == "000"], aes(color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()

```{r}
copy(miembros) %>% 
  .[, mean_RMSE := mean(RMSE), by =  .(exp, fcst_exp, date, var)] %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = fcst_exp, 
                grup = interaction(exp, fcst_exp, mem)), alpha = 0.1) +
  geom_line(aes(y = mean_RMSE, color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()
```
```{r}

extremos <- miembros %>% 
  .[, .(max_bias = max(BIAS),
        min_bias = min(BIAS)), by = .(var, exp, fcst_exp, date)]

media %>% 
  extremos[., on = .NATURAL] %>% 
  .[exp == "E7"] %>% 
  ggplot(aes(date, BIAS)) +
  geom_ribbon(aes(ymin = min_bias, 
                  ymax = max_bias, fill = exp, group = interaction(exp, fcst_exp)), alpha = 0.2) +
  scale_fill_manual(guide = NULL, values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                                   E6 = "SATWND", E7 = "RAD")) +
  geom_line(aes(color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()

media %>% 
  .[, mem := "000"] %>% 
  rbind(., miembros) %>% 
  .[exp == "E7"] %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(data = ~.x[mem != "000"], alpha = 0.1, 
            aes(y = RMSE, color = exp, linetype = fcst_exp, group = interaction(exp, mem, fcst_exp))) +
  geom_line(data = ~.x[mem == "000"], aes(color = exp, linetype = fcst_exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_wrap(~ var, scales = "free_y") +
  theme_minimal()

```{r}
cum <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
   setnames("cumulus", "param") %>% 
  .[, tipo := "cumulus"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., cum) %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

pbl <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(BIAS = mean(BIAS)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
   setnames("pbl", "param") %>% 
  .[, tipo := "pbl"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "BIAS")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(date, BIAS)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_date() +
  labs(subtitle = "BIAS", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

rbind(pbl, cum) %>%
  .[, .(BIAS_mean = mean(BIAS, na.rm = TRUE)), 
    by = .(var, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(BIAS_mean, exp)) +
  geom_vline(xintercept = 0, color = "grey70") +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ var, scales = "free_x") +
  labs(y = NULL, x = "mean BIAS", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))
```

```{r}
cum <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, cumulus, var, date)] %>% 
   setnames("cumulus", "param") %>% 
  .[, tipo := "cumulus"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., cum) %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(4, 2, 3, 1)) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

pbl <- fisica[miembros, on = .NATURAL] %>% 
  separate(fisica, into = c("cumulus", "pbl")) %>% 
  .[, .(RMSE = mean(RMSE)), by = .(exp, fcst_exp, pbl, var, date)] %>% 
   setnames("pbl", "param") %>% 
  .[, tipo := "pbl"]

media %>% 
  .[, c("exp", "var", "date", "fcst_exp", "RMSE")] %>% 
  .[, ":="(param = "Mean",
           tipo = "Mean")] %>% 
  rbind(., pbl) %>% 
  ggplot(aes(date, RMSE)) +
  geom_line(aes(color = exp, linetype = param)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                                      E6 = "SATWND", E7 = "RAD")) +
  # geom_hline(yintercept = 0, color = "grey30") +
  scale_linetype_manual(values = c(1, 2, 3, 4)) +
  scale_date() +
  labs(subtitle = "RMSE", x = NULL, y = NULL, linetype = NULL, color = NULL) +
  facet_grid(var~fcst_exp, scales = "free_y") +
  theme_minimal()

rbind(pbl, cum) %>%
  .[, .(RMSE_mean = mean(RMSE, na.rm = TRUE)), 
    by = .(var, exp, param, fcst_exp, tipo)] %>% 
  ggplot(aes(RMSE_mean, exp)) +
  geom_point(aes(color = exp, shape = param),
             size = 3) +
  scale_color_manual(guide = NULL, values = colores_exp, 
                      labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  scale_shape_manual(values = c(19, 17, 15, 3, 0, 8)) +
  scale_y_discrete(labels = c(E4 = "CONV", E5 = "AUT", 
                                 E6 = "SATWND", E7 = "RAD")) +
  facet_grid(fcst_exp ~ var, scales = "free_x") +
  labs(y = NULL, x = "mean RMSE", shape = NULL) +
  theme_minimal() +
  theme(aspect.ratio = 1/2,
        panel.background = element_rect(fill = "grey98",
                                        color = "grey98"))
```

### Precipitación

```{r}
files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_mean_acum_30h.rds"))

pp_acum <- purrr::map(files, function(f) {

  meta <- unglue(basename(f), "{exp}_fcst_{fcst_init}_mean_acum_30h.rds")

  read_rds(f) %>% 
    .[, pp_acum := if_else(pp_acum < 0, 0, pp_acum)] %>% 
    .[, ":="(exp = meta[[1]][["exp"]],
             fcst_init = meta[[1]][["fcst_init"]])]

}) %>% rbindlist()

pp_acum %>%

  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
                    proj = norargentina_lambert,
                    breaks = seq(20, 300, 20)) +
  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
                                                barheight = 0.8), 
                       super = ScaleDiscretised,
                       limits = c(20, 300),
                       na.value = "white") +
  geom_point(data = square, aes(lon, lat), size = 0.1) +
  geom_mapa() +
  facet_grid(fcst_init ~ exp, labeller = labeller(exp = c(E4 = "CONV", E5 = "AUT", 
                                                          E6 = "SATWND", E7 = "RAD"))) +
  labs(fill = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

  read_rds(paste0(derived_data, "ppacum/IMERG_30h.rds")) %>% 
  .[, run := "IMERG"] %>%
  .[,  c("lon", "lat") := wrf_project(x, y, inverse = TRUE, round = FALSE)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = pp_acum, fill = stat(level_d)), 
                    proj = norargentina_lambert,
                    breaks = seq(20, 300, 20)) +
  scale_fill_viridis_c(guide = guide_colorsteps(barwidth = 20,
                                                barheight = 0.8), 
                       super = ScaleDiscretised,
                       limits = c(20, 300),
                       na.value = "white") +
  geom_point(data = square, aes(lon, lat), size = 0.1) +
  geom_mapa() +
  facet_wrap(~run) +
  labs(fill = NULL) +
  theme_minimal() +
  theme(legend.position = "bottom") +

  plot_layout(ncol = 2, widths = c(3, 1), guides = "collect") & theme(legend.position = 'bottom')

```{r}
files <- Sys.glob(paste0(derived_data, "/ppacum/E[4-7]_fcst_201811220*_prop_acum_30h.rds"))

pp_prop <- purrr::map(files, function(f) {

  meta <- unglue(basename(f), "{exp}_fcst_{fcst_init}_prop_acum_30h.rds")

  read_rds(f) %>% 
    .[, ":="(exp = meta[[1]][["exp"]],
             fcst_init = meta[[1]][["fcst_init"]])]

}) %>% rbindlist()

>>>>

---
title: "Untitled"
output: html_document
---

@@

## Introduction

## Data and Methods

```{r dominio, fig.cap="The horizontal domain used on the simulations (black box) and the inner domain used for the validation / evaluation (grey box). The topography in meters is also shown (filled contours).", fig.height=4}

fread(here("analysis", "data", "derived_data", "dominio_hgt.csv")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>%

  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = hgt), proj = norargentina_lambert,
                    breaks = seq(0, 6000, 500)) +
  scale_fill_gradient(low = "#f2f2f2", high = "#333333",
                      name = NULL,
                       breaks = seq(0, 6000, 1000),
                       guide = guide_colorstrip(barwidth = 20,
                                                barheight = 0.8)) +
  geom_mapa() +
  geom_point(data = square, aes(lon, lat), size = 0.2) +
  geom_point(data = square2, aes(lon, lat), size = 0.2, color = "#CC3311") +
  theme_minimal() +
  theme(legend.position = "bottom", legend.box.spacing = margin(0, 0, 0, 0)) 
```
